Using Scipion for Stream Image Processing at EM Facilities

### 2018-03-09 0.abstract.tex

Three dimensional electron microscopy (3DEM) is becoming a very data-intensive field in which vast amounts of experimental images are acquired at high speed. To manage such large-scale projects, we have previously developed  a modular workflow system called \scipion \citep{delarosatrevin2016}. We present here a major extension of \scipion that allows stream processing of EM images while the data is being acquired. This approach helps to detect problems at early stages, saves computing time and provides users with an in deep evaluation of the data quality before the acquisition is finished. Different EM facilities have 
implemented their stream processing solution using custom-made scripts, but this script approach is not flexible and present a slow response time to new feature requests. The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc). This summary may be generated in HTML format and  published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided. At present, \scipion has been installed and is in production mode in several Cryo-EM facilities and it is being considered in many more.

>>>>

Three dimensional electron microscopy (3DEM) is becoming a very data-intensive field in which vast amounts of experimental images are acquired at high speed. To manage such large-scale projects, we have previously developed  a modular workflow system called \scipion \citep{delarosatrevin2016}. We present here a major extension of \scipion that allows stream processing of EM images while the data is being acquired. This approach helps to detect problems at early stages, saves computing time and provides users with an in deep evaluation of the data quality before the acquisition is finished. Different EM facilities have 
implemented their stream processing solution using custom-made scripts, but this script based approach is not flexible and present a slow response time to new feature requests. The advantages of our proposed solution are: (1) flexibility: users can customize their workflows from a  large collection of preloaded algorithms; (2) repeatability and traceability: workflows are stored and can be re-executed later, either with the same data or a different data set; (3) monitorization: there are utilities that constantly check how the execution of the algorithms is going and inform about possible anomalies; (4) reports: we have developed several graphical interfaces that are updated periodically producing a summary (e.g, CTF defocus values, system load, etc). This summary may be generated in HTML format and  published though a public web-site to provide access for external users; (5) backup: both local and remote backup of the data and image processing is provided. At present, \scipion has been installed and is in production mode in several Cryo-EM facilities and it is being considered in many more.

### 2018-03-09 1.introduction.tex

3D electron microscopy (3DEM)  can provide rich information about structural characteristics of macromolecular complexes. The field is under a major transformation due to the arrival of better microscopes, new large area detectors and automatization \citep{kuhlbrandt2014a, Kuhlbrandt2014b}  which makes possible that a single microscope generates high quality data sets on the order of terabytes per day \citep{Saibil2015} while working for several days without interruption. Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of a experienced microscopist and; (3) the need for an effective transference of data and associated metadata in an increasingly distributed and collaborative environment. This situation places techniques that are automated and allow higher throughput in high demand.

@@

Clearly, the future of projects that require high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff. \scipion has been redesigned with this scenario in mind in which the early detection of issues may save a data acquisition shift which otherwise needs to be repeated after a new application that may be granted many months later. In this way, \scipion starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and execute a first user-tailored image processing workflow that provides users with an accurate measurement of the acquired data quality.

Many software packages have been developed to provide automated data collection, such as SerialEM \citep{Mastronarde2005}, Leginon \citep{Suloway2009}, UCSF-Tomography \citep{Zheng2007}, Tom Toolbox \citep{Nickell2005}, EPU \citep{EPU}, Latitude in Digital Micrograph \citep{Latitude}, or EMMenu \citep{emmenu}, etc. However, the next step,  which is to provide image processing while  data are being collected, is not so well served. In fact, in most places this is accomplished through home made scripts (i.e. \citep{Pichkur2018}). Exceptions are UCSFImage4 \citep{Li2015} and Focus \citep{Biyani2017}. As we will show in this article, \scipion differs from these solutions in a larger flexibility both from the point of view of users and facility staff.

>>>>

3D electron microscopy (3DEM)  can provide rich information about structural characteristics of macromolecular complexes. The field is under a major transformation due to the arrival of better microscopes, new large area detectors and automatization \citep{kuhlbrandt2014a, Kuhlbrandt2014b}. These improvements make possible that a single microscope generates high quality data sets on the order of terabytes per day \citep{Saibil2015} while working for several days without interruption. Major challenges faced in the field are: (1) an efficient management of these huge datasets and their corresponding image processing workflows; (2) a raising interest from scientists outside the field who might lack the skills of a experienced microscopist and; (3) the need for an effective data and metadata transfer in an increasingly distributed and collaborative environment. This situation strongly demands the application of techniques that are highly automated to allow high throughput.

@@

Clearly, the future of projects that require high-end microscopes is through centralized microscopy facilities, in which automatic acquired data needs to be monitored by users and facility staff. [**JMRT VERY LARGE AND TWISTED SENTENCE \scipion has been redesigned with this scenario in mind in which the early detection of issues may save a data acquisition shift which otherwise needs to be repeated after a new application that may be granted many months later]. In this way, \scipion starts the processing of the movies as they are being acquired with the double aim of checking for possible data collection errors and execute a first user-tailored image processing workflow that provides users with an accurate measurement of the acquired data quality.

Many software packages have been developed to provide automated data collection, such as SerialEM \citep{Mastronarde2005}, Leginon \citep{Suloway2009}, UCSF-Tomography \citep{Zheng2007}, Tom Toolbox \citep{Nickell2005}, EPU \citep{EPU}, Latitude in Digital Micrograph \citep{Latitude}, or EMMenu \citep{emmenu}, etc. However, the next step,  which is to provide image processing while  data are being collected, is not so well addressed. In fact, in most places this is accomplished through home made scripts (i.e. \citep{Pichkur2018}). Exceptions are UCSFImage4 \citep{Li2015} and Focus \citep{Biyani2017}. As we will show in this article, \scipion differs from these solutions in a larger flexibility both from the point of view of users and facility staff.
